Spike spectra for recurrences

### 2022-07-21 Manuscript/inter_spike_manuscript_mdpi.tex
Merge branch 'main' of https://github.com/hkraemer/Recurrence_Spike_Spectra into main

\subsection{Period estimation for different dynamics in the R\"ossler system}
First, we consider the R\"ossler system (Eq.~\eqref{eq_model_roessler} in Appendix \ref{sec_models_roessler}) 
in three different dynamical setups. We use the proposed inter spike spectrum to
identify the type of dynamics.
We set the parameters $b=2$, $c=4$ and analyze period-2 limit cycle dynamics ($a=0.36$ in Fig.~\ref{fig_tau_rr_example_roessler}A, D, G, J), 
period-3 limit cycle dynamics ($a=0.41$ in Fig.~\ref{fig_tau_rr_example_roessler}B, E, H, K) and chaotic dynamics ($a=0.428$ in Fig.~\ref{fig_tau_rr_example_roessler}C, F, I, L).

\begin{figure}
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_example_roessler}
 \caption{Inter spike spectra of the $\tau$-RR of the R\"ossler system in three different dynamical regimes with parameters $b=2$, $c=4$. 
 \textbf{A} Trajectory of the system in a period-2 (parameter $a=0.36$), \textbf{B} in a period-3 (parameter $a=0.41$) and 
 \textbf{C} in a chaotic regime (parameter $a=0.428$). 
 \textbf{D, E, F} The corresponding RPs, obtained by using a recurrence threshold corresponding to a 10\% global 
 recurrence rate for D \& E and 5\% for F. 
  \textbf{G, H, I} $\tau$-RR's of the shown RPs. 
  \textbf{J, K, L} The proposed inter spike spectra of the $\tau$-RR's shown in panels G, H, I. Spectra were obtained with a LASSO regression and a regularization threshold 
  corresponding to $\rho=0.95$ accordance of $\tau$-RR's and re-composed signals. The distance ratio of the peaks reflect the limit cycle dynamic.  
  \textbf{M, N, P} Fourier power spectra of the $\tau$-RR's shown in panels G, H, I.
\todo[inline]{MA: why you have not used STLS method here?} }
\label{fig_tau_rr_example_roessler}
\end{figure}

@@

We consider the Logistic map $x_{n+1}=r\cdot x_n \left( 1-x_n \right)$ for changing control parameter $r$. We vary $r$ from $r=3.4$ to $r=4$ in steps of $0.001$. For 
each setting of $r$ 
\begin{itemize}[noitemsep]
\item[(1)] a time series of length $N=201$ is computed with a random initial condition $u_0 \in [0,\ 1]$, neglecting the first $1,000$ samples as transients,
\item[(2)] $100$ iterative Amplitude Adjusted Fourier Transform (iAAFT) surrogates \cite{Schreiber1996,Schreiber2000} are computed,
\item[(3)] the time series and its iAAFT surrogates are embedded in a 2-dimensional state space using a time delay of unity,
\item[(4)] from the 2-dimensional trajectories RPs, Eq.~\eqref{eq_rp_definition}, are computed under a threshold $\varepsilon=0.05$,
\item[(5)] $\tau$-RR, Eq.~\eqref{eq_tau_rr}, is computed from the RP of the signal and from the RPs of the surrogates,
\item[(6)] inter spike spectra are obtained from $\tau$-RR of the signal and from the $\tau$-RRs of the surrogates, see Section \ref{sec_tau_rr_method}, and finally,
\item[(7)] from the distribution of the surrogate inter spike spectra the $95^\text{th}$ percentile is computed. The peaks of the inter spike spectrum of the signal which exceed 
this percentile are counted. 
\end{itemize}
In this example, the according null hypothesis is that the data stems from a process which yields the same auto-correlation, hence the same Fourier powerspectrum, and the same 
amplitude distribution. We consider the number of significant peaks in the inter spike spectrum with
respect to the control parameter in order to distinguish the corresponding dynamics (Fig.~\ref{fig_tau_rr_logistic}C).
A correlation with the positive Lyapunov exponent (Fig.~\ref{fig_tau_rr_logistic}A) is discernible 
($\rho_{\text{Pearson}}=0.72$). Moreover, this analysis can tackle period-doubling, since it ``measures'' the dominant cycles via the inter spike spectrum.

@@

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{./figures/fig_tau_rr_logistic}
 \caption{\textbf{A} Lyapunov exponent of the Logistic map as a function of the control parameter $r$. 
 \textbf{B} Number of significant peaks ($\alpha=0.05$) in the inter spike spectrum of the $\tau$-RR and its Pearson correlation coefficient to the Lyapunov exponent shown in \textbf{A} 
 (white noise surrogates). 
 \textbf{C} Same as \textbf{B}, but for iterative Amplitude Adjusted Fourier Transform (iAAFT) surrogates \cite{Schreiber1996,Schreiber2000}. For obtaining the inter spike spectra we used 
 a LASSO regression and a regularization threshold corresponding to $\rho=0.95$ accordance of $\tau$-RR's and re-composed signals.  
}
\label{fig_tau_rr_logistic}
\end{figure}

>>>>

\subsection{Period estimation for different dynamics in the R\"ossler system}
First, we consider the R\"ossler system (Eq.~\eqref{eq_model_roessler} \todo{why not C.2?} Appendix \ref{sec_models_roessler}) 
in three different dynamical setups. We use the proposed inter spike spectrum to
identify the type of dynamics.
We set the parameters $b=2$, $c=4$ and analyze period-2 limit cycle dynamics ($a=0.36$ in Fig.~\ref{fig_tau_rr_example_roessler}A, D, G, J), 
period-3 limit cycle dynamics ($a=0.41$ in Fig.~\ref{fig_tau_rr_example_roessler}B, E, H, K) and chaotic dynamics ($a=0.428$ in Fig.~\ref{fig_tau_rr_example_roessler}C, F, I, L).

\begin{figure}
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_example_roessler}
 \caption{Inter spike spectra of the $\tau$-RR of the R\"ossler system in three different dynamical regimes with parameters $b=2$, $c=4$. 
 \textbf{A} Trajectory of the system in a period-2 (parameter $a=0.36$), \textbf{B} in a period-3 (parameter $a=0.41$) and 
 \textbf{C} in a chaotic regime (parameter $a=0.428$). 
 \textbf{D, E, F} The corresponding RPs, obtained by using a recurrence threshold corresponding to a 10\% global 
 recurrence rate for D \& E and 5\% for F \todo[inline]{what is the recurrence rate? does it relate to $\epsilon $ in Eq.(1).} 
  \textbf{G, H, I} $\tau$-RR's of the shown RPs. 
  \textbf{J, K, L} The proposed inter spike spectra of the $\tau$-RR's shown in panels G, H, I. Spectra were obtained with a LASSO regression and a regularization threshold 
  corresponding to $\rho=0.95$ accordance of $\tau$-RR's and re-composed signals. The distance ratio of the peaks reflect the limit cycle dynamic.  
  \textbf{M, N, P} Fourier power spectra of the $\tau$-RR's shown in panels G, H, I.
\todo[inline]{MA: why you have not used STLS method here?} }
\label{fig_tau_rr_example_roessler}
\end{figure}

@@

We consider the Logistic map $x_{n+1}=r\cdot x_n \left( 1-x_n \right)$ for changing control parameter $r$. We vary $r$ from $r=3.4$ to $r=4$ in steps of $0.001$ \todo{is it correct?}. For 
each setting of $r$ 
\begin{itemize}[noitemsep]
\item[(1)] a time series of length $N=201$ is computed with a random initial condition $u_0 \in [0,\ 1]$, neglecting the first $1,000$ samples as transients,
\item[(2)] $100$ iterative Amplitude Adjusted Fourier Transform (iAAFT) surrogates \cite{Schreiber1996,Schreiber2000} are computed,
\item[(3)] the time series and its iAAFT surrogates are embedded in a 2-dimensional state space using a time delay of unity,
\item[(4)] from the 2-dimensional trajectories RPs, Eq.~\eqref{eq_rp_definition}, are computed under a threshold $\varepsilon=0.05$,
\item[(5)] $\tau$-RR, Eq.~\eqref{eq_tau_rr}, is computed from the RP of the signal and from the RPs of the surrogates,
\item[(6)] inter spike spectra are obtained from $\tau$-RR of the signal and from the $\tau$-RRs of the surrogates, see Section \ref{sec_tau_rr_method}, and finally,
\item[(7)] from the distribution of the surrogate inter spike spectra the $95^\text{th}$ percentile is computed. The peaks of the inter spike spectrum of the signal which exceed 
this percentile are counted. 
\end{itemize}
In this example, the according null hypothesis \todo{null hypothesis refers to surrogate data?}is that the data stems from a process which yields the same auto-correlation, hence the same Fourier powerspectrum, and the same 
amplitude distribution. We consider the number of significant peaks in the inter spike spectrum with
respect to the control parameter in order to distinguish the corresponding dynamics (Fig.~\ref{fig_tau_rr_logistic}C).
A correlation with the positive Lyapunov exponent (Fig.~\ref{fig_tau_rr_logistic}A) is discernible 
($\rho_{\text{Pearson}}=0.72$). Moreover, this analysis can tackle period-doubling, since it ``measures'' the dominant cycles via the inter spike spectrum. \todo{This part is not understandable}

@@

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{./figures/fig_tau_rr_logistic}
 \caption{\textbf{A} Lyapunov exponent of the Logistic map as a function of the control parameter $r$. \todo[inline]{A shows also bifurcation diagram?}
 \textbf{B} Number of significant peaks ($\alpha=0.05$) in the inter spike spectrum of the $\tau$-RR and its Pearson correlation coefficient to the Lyapunov exponent shown in \textbf{A} 
 (white noise surrogates). 
 \textbf{C} Same as \textbf{B}, but for iterative Amplitude Adjusted Fourier Transform (iAAFT) surrogates \cite{Schreiber1996,Schreiber2000}. For obtaining the inter spike spectra we used 
 a LASSO regression and a regularization threshold corresponding to $\rho=0.95$ accordance of $\tau$-RR's and re-composed signals.  
}
\label{fig_tau_rr_logistic}
\end{figure}
